---
title: "assignment1_db"
author: "Dongbo Shi"
date: "5/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Peer Assignment 1 of Reproducible Research

## Load data 
```{r load data, echo=TRUE}
library(readr)
activity <- read_csv("/Users/birdstone/Downloads/activity.csv")
```

## Histogram of the total number of steps taken each day
```{r total number, echo=TRUE}
library(dplyr)
activity_hist <- activity %>% filter(!is.na(steps)) %>% group_by(date) %>%
        summarise(total_steps=sum(steps))
hist(activity_hist$total_steps)
```

## Mean and median number of steps taken each day
```{r mean, echo = TRUE}
mean(activity$steps,na.rm = T)
median(activity$steps,na.rm = T)
```

## Time series plot of the average number of steps taken
```{r average plot, echo = TRUE}
activity_plot <- activity %>% filter(!is.na(steps)) %>% group_by(date) %>%
        summarise(average_steps=mean(steps))
plot(activity_plot$date,activity_plot$average_steps,type = "l",xlab = "Date",ylab = "Average Steps")
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r max step, echo=TRUE}
activity_naomit <- activity %>% filter(!is.na(steps))
activity_naomit$interval[activity_naomit$steps == max(activity_naomit$steps,na.rm = T)]
```

## Code to describe and show a strategy for imputing missing data
## Histogram of the total number of steps taken each day after missing values are imputed
```{r imputing, echo=TRUE}
# the imputing strategy is replace the missing data with mean steps
activity_impute <- activity %>% mutate(steps = if_else(is.na(steps),mean(activity$steps,na.rm = T),steps))
activity_impt_hist <- activity_impute %>% group_by(date) %>%
        summarise(total_steps=sum(steps))
hist(activity_impt_hist$total_steps)
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r comparison, echo=TRUE}
activity_impute <- activity_impute %>% mutate(weekends=if_else(weekdays(activity_impute$date) %in% c("Saturday","Sunday"),"weekend","weekday")) 
activity_impt_plot <- activity_impute %>% group_by(weekends,interval) %>% summarise(average_steps=mean(steps))
library(ggplot2)
p1 <- ggplot(data = activity_impt_plot,aes(x = interval, y = average_steps)) +
        geom_line() + facet_wrap(~weekends,ncol=1) + ylab("Average Steps")
p1
```
